Target Speaker Extraction using Discrete Representations from Self-Supervised Models and Language Models \thanks{Identify applicable funding agency here. If none, delete this.}

### 2024-09-04 src/paper.tex
update

\subsection{Encoding}
We use pretrained SSL model WavLM Large \cite{wavlm} to 
encode the speech into continuous representations. We use the output from 6 hidden 
layers 1, 3, 7, 12, 18, 23. Given a speech signal \(s \in R^{T'} \), the output 
of WavLM will be a tensor \(\bm{r}\) of shape \(n \times T \times E\) where 
\(n\) specifies the number 
of output layers (6 in this case), and  \(T\) is the time 
dimension and \(E\) stands for the 
embedding dimension, specifically 1024 in WavLM Large. The Kmeans tokenizer uses \(n\) numbers of  Kmeans model 
separately on each output layer. Each model uses the same clustering number denoted by \(K\).
After tokenization, the continuous embedding \(\bm{r}\) will be transformed to a tensor \(\bm{d}\) of shape \(n \times T\) where each value \(\bm{d}_{(n_i,t)} \in (0, K-1), n_i \in (0,n), t \in 
(0, T)\). We choose \(K\) to be 1000 in our experiments. For reference speech and mixed speech, we 
use the same Kmeans model and the same layer output of WavLM.

The encoding strategy on the mixed speech plays a pivotal role in the model performance. Assume we 
have a reference speech \(s_r \in R^{T^r} \) and mixed speech \(s \in R^{T'}\). 
We follow the 
aforementioned procedure for 
reference speech to get tensor \(\bm{d_r}\) of shape \(n \times T_r\). But for the mixed 
speech, instead of directly following the procedure, we first concatenate it with 
the reference speech to get \(s' = [s_r, s_m, s_r] \in R^{(T^r+T'+T^r)}\). Then we use this speech 
as input for the encoder, obtaining output \(\bm{d'}\) of shape \(n \times (T_r+T+T_r)\) where \(T\) 
denotes the length of output by directly passing the mixed speech without concatenation.

>>>>

\subsection{Encoding}
We use pretrained SSL model WavLM Large \cite{wavlm} to 
encode the speech into continuous representations. We use the output from 6 hidden 
layers 1, 3, 7, 12, 18, 23. Given a speech signal \(s \in R^{T'} \), the output 
of WavLM will be a tensor \(\bm{r}\) of shape \(n \times T \times E\) where 
\(n\) specifies the number 
of output layers (6 in this case), and  \(T\) is the time 
dimension and \(E\) stands for the 
embedding dimension, specifically 1024 in WavLM Large. The Kmeans tokenizer uses \(n\) numbers of  Kmeans model 
separately on each output layer. Each model uses the same clustering number denoted by \(K\).
After tokenization, the continuous embedding \(\bm{r}\) will be transformed to a tensor \(\bm{d}\) of shape \(n \times T\) where each value \(\bm{d}_{(n_i,t)} \in (0, K-1), n_i \in (0,n), t \in 
(0, T)\). We choose \(K\) to be 1000 in our experiments. For reference speech and mixed speech, we 
use the same Kmeans model and the same layer output of WavLM. The encoder is freezed during training.

The encoding strategy on the mixed speech plays a pivotal role in the model performance. Assume we 
have a reference speech \(s_r \in R^{T^r} \) and mixed speech \(s \in R^{T'}\). 
We follow the 
aforementioned procedure for 
reference speech to get tensor \(\bm{d_r}\) of shape \(n \times T_r\). But for the mixed 
speech, instead of directly following the procedure, we first concatenate it with 
the reference speech to get \(s' = [s_r, s_m, s_r] \in R^{(T^r+T'+T^r)}\). Then we use this speech 
as input for the encoder, obtaining output \(\bm{d'}\) of shape \(n \times (T_r+T+T_r)\) where \(T\) 
denotes the length of output by directly passing the mixed speech without concatenation. \(\bm{d'}\) 
contains the discrete tokens for two pieces of the same reference speech and one piece of the mixed 
speech, we select tensor \(\bm{d}\) belonging only to the mixed speech of shape \(n\times T\) as the 
encoder output. This process is inspired by the training schema of WavLM \cite{wavlm}. To ensure the 
model correctly outputs rich information of the target speaker, it is trained by 
having a clean speech overlapped with another interference less than 50\% of the length of the clean 
speech and using the first utterance as the main speaker. By this way, the model identifiers the 
main speaker and produces embedding focusing on the main speaker. Our later experiments showed that 
this strategy plays an important role in the model performance.

\subsection{Modeling}
